---
title: "CORI microplastic data analysis workflow"
author: "Ania"
date: "January 23, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hello! 

> ####Familiarize yourself with the dataset of the CORI lab
####Naming conventions, important:

* no typos [but we will check]
* as short as possible & meaningful [for plotting we can change]
* no spaces, and best to seperate with *_* , for e.g. filter_blank
  
  
  
  
> ####Variables in the input data sheet, their meaning and what are they used for

* **project** [char] | enter your porject ID
* **site** [char] | enter here your site ID
* **flag** [int] | enter a numeric (1,2,3,...) that means something to you
* **sample** [char] | combo of numbers and letters, e.g. M1 (mussel 1)
* **date_extracted** | [ddmmyy], for folks who do digestions
* **date_inspected** | [ddmmyy], hmm?
* **photo_number** | numeric (1,2,3) or ID?
* **proc_blank** | not sure if this is for a seperate spreadsheet to be joined
* **bckg_blank** | not sure if this is for a seperate spreadsheet to be joined
* **mp_number** | a particle you suspect is microplastic that you count on the filter. This is SUSPECTED microplastic
* **mp_color** | describes the color of your MP [blue, red, pink, organge]
* **mp_type** | fiber, sheet, particle - EVERYONE MUST USE CONSISTENT NAMES!
* **short_axis** | in microns, microplastic bit width
* **long_axis** | in microns, microlastic length
* **FTIR_conf** | binary : Y or N [NOTE this must be capitalized!], if you choose Y - your suspected MP is actual MP as confirmed by the FITR, if ou choose N - your suspected MP is NOT and MP
* **MP_material** | [character] polyester, nylon etc - EVERYONE MUST USE CONSISTENT NAMES!


> ####Variables in the output data sheet #1 (csv called "Total_mp-per-sample ") and their meaning:
  
* **total_mp_number** [int] | this is the sum of **mp_number** per sample; total of the suspected MPs
* **corr_mp_number** [int] | this is a sum of your MPs per sample once they have been confirmed with FTIR. To obtain this, samples that have Y in the **FTIR_conf** column are filtered and summarized (in other words we exclude particles that are NOT microplastics)
*

> ####Variables in the output data sheet # 2 (csv called "Percent_mp-type-per-sample") and their meaning:

* table contains sample ID and % 

> ####Steps:

1. Subtract the blank - assess background & procedural blank: how to?
2. Summarize total microplastic per sample + report
3. Remove particles that are not microplastics -> summarize -> report
4. How do people want to have TYPE, COLOR and MATERIAL reported? / visualized?


> Preliminary processing

##### Taking a look at a mock dataset from Julie

```{r, message=FALSE}

library(tidyverse)
mp_data <- read.csv("R_coding/Microplastics_data-workflow-OPRP/MP_demo-dataset.csv")

glimpse(mp_data)
#knitr::kable(head(mp_data), format = "html")

```

### Summarizing total microplastics (includes suspected). The first command generates a table, which can be written to csv. ** If you put the whole command in bracket it will print

```{r, message=FALSE}

(summary <- mp_data %>%
  group_by(site,sample) %>% # this means: calculate a sum for every station and every sample
  summarize(sum(mp_number))) # function that does the sum

# write.csv( summary, "Total_suspected-microplastics.csv") - this is your saved data

```

we can implement the above commands into a plot - this is good when exploring things

```{r, message=FALSE}

mp_data %>%
  group_by(site,sample) %>% # this means: calculate a sum for every station and every sample
  summarize(mp_susp = sum(mp_number)) %>% 
  ggplot(aes(site, mp_susp, fill = site))+
  geom_bar(stat = "identity")# function that does the sum

# write.csv( summary, "Total_suspected-microplastics.csv") - this is your saved data

```
> 

#### Per each site and sample let's make them plots (yoda)
```{r, message=FALSE}

mp_data %>%
  group_by(site,sample) %>% # this means: calculate a sum for every station and every sample
  summarize(mp_susp = sum(mp_number)) %>% 
  ggplot(aes(sample, mp_susp, fill = site))+
  geom_bar(stat = "identity") +
  facet_grid (~ site)

```

> 

#### What is the distribution of MP type in each sample### Let's see what type of MP we have in each sample



